Component response rate variation underlies the stability of complex systems (revision notes)

### 2020-01-18 notebook/revision_notes.Rmd
Minor changes

Given that diagonal values of $\textbf{M}$ are set to $-1$, We can know the relationship $S = 1600$, $C = 1$, and $\sigma = 1/40$. Under these parameter values, given $\rho = 0$, we know that $\sigma\sqrt{SC}\left(1 + \rho\right) = 1$. Hence, the expected value of $\max(\Re(\lambda)) = 0$

@@

Hence, under the same set of parameters, we predict a linear effect of $\rho$ on $\max(\Re(\lambda)) = 0$, such that $\max(\Re(\lambda)) = \rho$.

Allesina and Tang note that $\rho$ is defined as follows [@Allesina2015a],

@@

We can rewrite as the more typical way of writing covariance (top),

>>>>

Given that diagonal values of $\textbf{M}$ are set to $-1$, we can consider the relationship $S = 1600$, $C = 1$, and $\sigma = 1/40$. Under these parameter values, given $\rho = 0$, we know that $\sigma\sqrt{SC}\left(1 + \rho\right) = 1$. Hence, the expected value of $\max(\Re(\lambda)) = 0$

@@

Hence, under the same set of parameters, we predict a linear effect of $\rho$ on $\max(\Re(\lambda)) = 0$, such that $\max(\Re(\lambda)) = \rho$. Allesina and Tang note that $\rho$ is defined as follows [@Allesina2015a],

@@

Since we define $\sigma = \sqrt{S}$ and $C = 1$, $\sigma\sqrt{SC} = 1$, so the distribution of eigenvalues is circular with a radius of unity. Further, since we set $diag(\textbf{M}) = -1$, we know that $E[\max(\Re(\lambda))] = 0$.

Gibbs et al. [@Gibbs2017] showed analytically that as $S \to \infty$, the effect of a diagonal matrix $\textbf{X}$ on the stability of $\textbf{M} = \textbf{XA}$ becomes negligible. Hence, assuming high $S$ and $\rho = 0$, we should not see an effect of $\gamma$ on $E[\max(\Re(\lambda))]$. But there is clearly an effect of $\gamma$ for nonzero $\rho$ such that $\partial E[\max(\Re(\lambda))] / \partial\rho > 0$ and $\partial^{2} E[\max(\Re(\lambda))] / \partial\rho > 0$.

Note that, by definition, we have not changed $E[M_{ij}]$ with the addition of $\gamma$, since $E[\gamma_{ii}] = 1$. Elements of $\gamma$ are pulled from a random uniform distribution between 0 and 2, meaning that no covariance should be induced betwen $M_{ij}$ and $M_{ji}$.

### 2020-01-18 notebook/revision_notes.tex
Minor changes

\hypertarget{role-of-correlated-matrices-in-stabilisation}{%
\section{Role of correlated matrices in
stabilisation}\label{role-of-correlated-matrices-in-stabilisation}}

@@

When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
\(\rho\) was normally distributed around
\ensuremath{4.8394472\times 10^{-6}} (95\% CIs:
\ensuremath{-7.3124139\times 10^{-5}},
\ensuremath{8.7487968\times 10^{-5}}). Hence, decreasing the correlation
between \(\textbf{A}_{ij}\) and \(\textbf{A}_{ji}\) was not by itself
the cause of stability. For a clearer picture of the effect of
\(\gamma\), it is useful to show the relationship between \(\rho\) and
\(\max(\Re(\lambda))\) again as above, but this time also for how the
relationship changes given \(Var(\gamma)\). That is, given
\(\textbf{M} = \textbf{A}\) in addition to
\(\textbf{M} = \gamma\textbf{A}\).

@@

\hypertarget{trying-to-think-analytically}{%
\section{Trying to think
analytically}\label{trying-to-think-analytically}}

@@

Given that diagonal values of \(\textbf{M}\) are set to \(-1\), We can
know the relationship \(S = 1600\), \(C = 1\), and \(\sigma = 1/40\).
Under these parameter values, given \(\rho = 0\), we know that
\(\sigma\sqrt{SC}\left(1 + \rho\right) = 1\). Hence, the expected value
of \(\max(\Re(\lambda)) = 0\)

@@

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Allesina2015a}{}%
1. Allesina, S. \& Tang, S. The stability--complexity relationship at
age 40: a random matrix perspective. \emph{Population Ecology} 63--75
(2015).
doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}

>>>>

\section{Role of correlated matrices in
stabilisation}\label{role-of-correlated-matrices-in-stabilisation}

@@

When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
\(\rho\) was normally distributed around 4.8394472\times 10\^{}\{-6\}
(95\% CIs: -7.0762643\times 10\^{}\{-5\}, 8.0991595\times 10\^{}\{-5\}).
Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
\(\textbf{A}_{ji}\) was not by itself the cause of stability. For a
clearer picture of the effect of \(\gamma\), it is useful to show the
relationship between \(\rho\) and \(\max(\Re(\lambda))\) again as above,
but this time also for how the relationship changes given
\(Var(\gamma)\). That is, given \(\textbf{M} = \textbf{A}\) in addition
to \(\textbf{M} = \gamma\textbf{A}\).

@@

\section{Trying to think
analytically}\label{trying-to-think-analytically}

@@

Given that diagonal values of \(\textbf{M}\) are set to \(-1\), we can
consider the relationship \(S = 1600\), \(C = 1\), and
\(\sigma = 1/40\). Under these parameter values, given \(\rho = 0\), we
know that \(\sigma\sqrt{SC}\left(1 + \rho\right) = 1\). Hence, the
expected value of \(\max(\Re(\lambda)) = 0\)

@@

Hence, under the same set of parameters, we predict a linear effect of
\(\rho\) on \(\max(\Re(\lambda)) = 0\), such that
\(\max(\Re(\lambda)) = \rho\).

Allesina and Tang note that \(\rho\) is defined as
follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{1}},

\[\rho = \frac{E[M_{ij}M_{ji}] - E[M_{ij}]E[M_{ji}]}{Var[M_{ij}]}.\]

We can rewrite as the more typical way of writing covariance (top),

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Allesina2015a}{}
1. Allesina, S. \& Tang, S. The stability--complexity relationship at
age 40: a random matrix perspective. \emph{Population Ecology} 63--75
(2015).
doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}
